---
title: "Tidy data and the grammar of graphics"
---

## Objectives 

- Learning the `facet_()` function and its usage
- Practicing creating some group-by-summary plots using `ggplot2`

## Setup

- Check your `dplyr` package is up-to-date by typing `packageVersion("dplyr")`. If the current installed version is less than 1.0, then update by typing `update.packages("dplyr")`. You may need to restart R to make it work.

```{r}
ifelse(packageVersion("dplyr") >= 1,
  "The installed version of dplyr package is greater than or equal to 1.0.0", update.packages("dplyr")
)

if (!require("pacman")) install.packages("pacman")

pacman::p_load(
  tidyverse, # the tidyverse framework
  here, # computational reproducibility
  gapminder, # toy data
  ggthemes, # additional themes
  ggrepel, # arranging ggplots
  patchwork, # arranging ggplots
  broom # tidying model outputs
)
```

### Many plots 

Basic ideas:

- Grouping: tell `ggplot2` about the structure of your data 
- Facetting: break up your data into pieces for a plot 

#### Grouping

- Can you guess what's wrong?

```{r}
p <- ggplot(gapminder, aes(x = year, y = gdpPercap))

p + geom_point()

p + geom_line()
```

`geom_line` joins up all the lines for each particular year in the order they appear in the dataset. `ggplot2` does not know the yearly observations in your data are grouped by country. 

You need grouping when the grouping information you need to tell is not built into the mapped variables (like continent).

```{r}
gapminder
```

#### Facetting 

Facetting is to make small multiples. 

- `facet_wrap`: based on a single categorical variable like `facet_wrap(~single_categorical_variable)`. Your panels will be laid out in order and then wrapped into a grid.

- `facet_grid`: when you want to cross-classify some data by two categorical variables like `facet_grid(one_cat_variable ~ two_cat_variable)`. 

```{r}
p <- ggplot(gapminder, aes(x = year, y = gdpPercap))

p + geom_line(aes(group = country)) # group by, # The outlier is Kuwait.
```

```{r}
p + geom_line(aes(group = country)) + facet_wrap(~continent) # facetting
```

```{r}
p + geom_line(aes(group = country), color = "gray70") +
  geom_smooth(size = 1.1, method = "loess", se = FALSE) +
  scale_y_log10(labels = scales::dollar) +
  facet_wrap(~continent, ncol = 5) + # for single categorical variable; for multiple categorical variables use facet_grid()
  labs(
    x = "Year",
    y = "GDP per capita",
    title = "GDP per capita on Five continents"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
p + geom_line(aes(group = country), color = "gray70") +
  geom_smooth(size = 1.1, method = "loess", se = FALSE) +
  scale_y_log10(labels = scales::dollar) +
  facet_grid(~continent) + # for single categorical variable; for multiple categorical variables use facet_grid()
  labs(
    x = "Year",
    y = "GDP per capita",
    title = "GDP per capita on Five continents"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


### Transforming

- Transforming: perform some calculations on or summarize your data before producing the plot 

#### Use pipes to summarize data

Also, we experiment with bar charts here. By default, `geom_bar` [uses](https://www.rdocumentation.org/packages/ggplot2/versions/1.0.1/topics/geom_bar) stat = "bins", which makes the height of each bar equal to the number of cases in each group. If you have a y column, then you should use `stat = "identity"` argument. Alternatively, you can use `geom_col().`

```{r}

gapminder_formatted <- gapminder %>%
  group_by(continent, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  )

ggplot(data = gapminder_formatted, aes(x = year, y = lifeExp_mean, color = continent)) +
  geom_point() +
  labs(
    x = "Year",
    y = "Life expectancy",
    title = "Life expectancy on Five continents"
  )
```

```{r}
gapminder %>%
  dplyr::filter(continent == "Europe") %>%
  group_by(country, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = year, y = lifeExp_mean, color = country)) +
  geom_point() +
  labs(
    x = "Year",
    y = "Life expectancy",
    title = "Life expectancy in Europe"
  )
```

```{r}
# geom point
gapminder %>%
  dplyr::filter(continent == "Europe") %>%
  group_by(country, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = year, y = lifeExp_mean)) +
  geom_point() +
  labs(
    x = "Year",
    y = "Life expectancy",
    title = "Life expectancy in Europe"
  ) +
  facet_wrap(~country)
```

```{r}
# geom bar
gapminder %>%
  dplyr::filter(continent == "Europe") %>%
  group_by(country, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = year, y = lifeExp_mean)) +
  geom_bar(stat = "identity") +
  labs(
    x = "Year",
    y = "Life expectancy",
    title = "Life expectancy in Europe"
  ) +
  facet_wrap(~country)
```

```{r}
# no facet
gapminder %>%
  dplyr::filter(continent == "Europe") %>%
  group_by(country, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = year, y = lifeExp_mean, fill = country)) +
  geom_bar(stat = "identity") + # even if you not stack, still the plot looks messy or you can use geom_col()
  labs(
    x = "Year",
    y = "Life expectancy",
    title = "Life expectancy in Europe"
  )
```

```{r}
gapminder %>%
  dplyr::filter(continent == "Europe") %>%
  group_by(country, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = country, y = lifeExp_mean)) +
  geom_boxplot() +
  labs(
    x = "Country",
    y = "Life expectancy",
    title = "Life expectancy in Europe"
  ) +
  coord_flip()
```

```{r}
# without ordering
gapminder %>%
  dplyr::filter(continent == "Europe") %>%
  group_by(country, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = reorder(country, lifeExp_mean), y = lifeExp_mean)) +
  geom_boxplot() +
  labs(
    x = "Country",
    y = "Life expectancy",
    title = "Life expectancy in Europe"
  ) +
  coord_flip()

# reorder
gapminder %>%
  dplyr::filter(continent == "Europe") %>%
  group_by(country, year) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = reorder(country, -lifeExp_mean), y = lifeExp_mean)) +
  geom_boxplot() +
  labs(
    x = "Country",
    y = "Life expectancy",
    title = "Life expectancy in Europe"
  ) +
  coord_flip()
```

#### Plotting text

```{r}
gapminder %>%
  dplyr::filter(continent == "Asia" | continent == "Americas") %>%
  group_by(continent, country) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = gdp_mean, y = lifeExp_mean)) +
  geom_point() +
  geom_text(aes(label = country)) +
  scale_x_log10() +
  facet_grid(~continent)
```

```{r}
# with label
gapminder %>%
  dplyr::filter(continent == "Asia" | continent == "Americas") %>%
  group_by(continent, country) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = gdp_mean, y = lifeExp_mean)) +
  geom_point() +
  geom_label(aes(label = country)) +
  scale_x_log10() +
  facet_grid(~continent)
```

```{r}
# no overlaps
gapminder %>%
  dplyr::filter(continent == "Asia" | continent == "Americas") %>%
  group_by(continent, country) %>%
  summarize(
    gdp_mean = mean(gdpPercap),
    lifeExp_mean = mean(lifeExp)
  ) %>%
  ggplot(aes(x = gdp_mean, y = lifeExp_mean)) +
  geom_point() +
  geom_text_repel(aes(label = country)) + # there's also geom_label_repel
  scale_x_log10() +
  facet_grid(~continent)
```
