---
title: "Assignment2"
author: "<Insert your name please>"
output: html_document
date: "`r Sys.Date()`"
---

# Assignment 2: Programming

1. Install the COVID-19 data package in R: `install.packages("COVID19")`

Documentation is here: https://covid19datahub.io/articles/r.html 

```{r}
# Install packages 
if(!require(pacman)) install.packages("pacman")

pacman::p_load(tidyverse, tidylog, here, COVID19)
```

2. Read the FT article on ["Covid-19 vaccine tracker: the global race to vaccinate."](https://ig.ft.com/coronavirus-vaccine-tracker/?areas=gbr&areas=isr&areas=usa&areas=eue&areas=are&areas=chn&areas=chl&cumulative=1&doses=total&populationAdjusted=1)

3. *Programming*: Visualize how the vaccination rate varies at the subnational level (state or even lower) in a particular country of your choice. [point: 50]

```{r}
# state level data 
# df <- covid19(country = c("US"), level = 2) # The US 
```

4. *Information*: Explain what kind of statistical elements are included in your plot(s). [point: 20]

5. *Communication*: Describe what kind of narrative your visualization conveys to which segment of your target audience (define this too) and why you expect it would work. [point: 30]

## Grading 

* Check - (0-50 points: 50 points)

* Check (51-70 points: 70 points) 

* Check + (71-100 points: 100 points)
