# Mapping

## Objectives 

- Learning ... tktk 
- Practicing ... tktk

## Setup

- Check your `dplyr` package is up-to-date by typing `packageVersion("dplyr")`. If the current installed version is less than 1.0, then update by typing `update.packages("dplyr")`. You may need to restart R to make it work.

```{r}
ifelse(packageVersion("dplyr") >= 1,
  "The installed version of dplyr package is greater than or equal to 1.0.0", update.packages("dplyr")
)

if (!require("pacman")) install.packages("pacman")

pacman::p_load(
  tidyverse, # the tidyverse framework
  here, # computational reproducibility
  gapminder, # toy data
  ggthemes, # additional themes
  ggrepel, # arranging ggplots
  patchwork, # arranging ggplots
  broom # tidying model outputs
)
```

## Ploting models 

In plotting models, we extensively use David Robinson's [broom package](https://cran.r-project.org/web/packages/broom/vignettes/broom.html) in R. The idea is to transform model outputs (i.e., predictions and estimations) into tidy objects so that we can easily combine, separate, and visualize these elements. 

### Plotting several fits at the same time

```{r}
model_colors <- RColorBrewer::brewer.pal(3, "Set1") # select three qualitatively different colors from a larger palette.

gapminder %>%
  ggplot(aes(x = log(gdpPercap), y = lifeExp)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", aes(color = "OLS", fill = "OLS")) +
  geom_smooth(
    method = "lm", formula = y ~ splines::bs(x, df = 3),
    aes(color = "Cubic Spline", fill = "Cubic Spline")
  ) +
  geom_smooth(method = "loess", aes(color = "LOESS", fill = "LOESS")) +
  theme(legend.position = "top") +
  scale_color_manual(name = "Models", values = model_colors) +
  scale_fill_manual(name = "Models", values = model_colors)
```

### Extracting model outcomes 

```{r}

# regression model
out <- lm(
  formula = lifeExp ~ gdpPercap + pop + continent,
  data = gapminder
)
```

`tidy()` is a method in the `broom` package. It "constructs a dataframe that summarizes the model's statistical findings". As the description states, tidy is a function that can be used for various models. For instance, a tidy can extract the following information from a regression model.

- `Term`: a term being estimated 
- `p.value`
- `statistic`: a test statistic used to compute p-value
- `estimate` 
- `conf.low`: the low end of a confidence interval 
- `conf.high`: the high end of a confidence interval
- `df`: degrees of freedom

**Challenge**

Try `glance(out)`; what did you get from these commands? If you're curious, you can try `?glance`.

The followings are to show your degree of confidence.

#### Coefficients

```{r}
# estimates
out_comp <- tidy(out)

p <- out_comp %>%
  ggplot(aes(x = term, y = estimate))

p + geom_point() +
  coord_flip() +
  theme_bw()
```

#### Confidence intervals

```{r}
# plus confidence intervals
out_conf <- tidy(out, conf.int = TRUE)

# plotting coefficients using ggplot2 (pointrange)
out_conf %>%
  ggplot(aes(x = reorder(term, estimate), y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  coord_flip() +
  labs(x = "", y = "OLS Estimate") +
  theme_bw()

# another way to do it (errorbar)
out_conf %>%
  ggplot(aes(x = estimate, y = reorder(term, estimate))) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
  labs(y = "", x = "OLS Estimate") +
  theme_bw()
```

You can also calculate marginal effects using the [`marginaleffects`](https://vincentarelbundock.github.io/marginaleffects/) package. 
